---
title: "Add Health Data Preparation for G-E Interplay on EXT Trajectories"
author: "LiChen Dong"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---


## Setup

```{r}
#| output: false

options(conflicts.policy = "depends.ok", scipen = 999, pillar.max_dec_width = 999)
library(tidyverse)
library(lavaan)

path_data <- "../data_local"
```


## Read in data 

### Add health in-home interview data

```{r}
load(here::here(path_data, "add_health_core.RData"))
```

### Genetic data ID - Phenotypic data ID linking file

```{r}
gid <- data.table::fread(here::here(path_data, "AddHealth dbGap ID Link File.dat")) |> as_tibble() |> mutate(AID = as.character(AID))
```

### Genetic ancestry

```{r}
ancestry <- foreign::read.xport(here::here(path_data, "PGS_AH1.xpt")) |> select(AID, ANCESTRY) |>  mutate(AID = as.character(AID))

```

### Polygenic score and genetic PCs

```{r}
ext_pgs <- read_csv(here::here(path_data, "8.14.24_EXT_PGS_PRSCS.csv")) |>
          mutate(EXT_PGS = scale(EXT_PGS_PRSCS) |> as.vector(),
                 AID = as.character(AID)) |> 
          select(-EXT_PGS_PRSCS)

genetic_pc <- data.table::fread(here::here(path_data, "ADDHPCA"))
names(genetic_pc) <- c("GFAM","GID",paste0("PC",1:20))

pgs_pc <- left_join(gid, ext_pgs) |>
           left_join(ancestry) |> 
           left_join(genetic_pc) |> 
           mutate(AID = as.character(AID))
```

## Creating variables

### AID

```{r}
data <- Wave1 |> select(AID)
```

### Participant Biological sex

```{r}
sex <- Wave1 |> select(AID, BIO_SEX) |> 
       rename(sex = BIO_SEX) |> 
       mutate(sex = replace(sex, sex > 5, NA),
              sex = case_match(sex, 1 ~ -.5, 2 ~ .5))
```

### Age

```{r}
age <- Wave1 |> select(AID, H1GI1M, H1GI1Y, IMONTH, IYEAR) |> 
                mutate(H1GI1M = replace(H1GI1M, H1GI1M == 96, NA), 
                       H1GI1Y = replace(H1GI1Y, H1GI1Y == 96, NA)) |> 
                mutate(age = IYEAR - H1GI1Y + (IMONTH - H1GI1M)/12) |> 
                select(-H1GI1M, -H1GI1Y, -IMONTH, -IYEAR) 

```

### Race/ethnicity

```{r}
race <- Wave1 |> select(AID, H1GI6A, H1GI6B, H1GI6C, H1GI6D, H1GI6E, H1GI4) |> 
        mutate(across(H1GI6A:H1GI4, ~ replace(.x, .x > 5, NA))) |> 
        rename(white = H1GI6A, african = H1GI6B, native = H1GI6C, asian = H1GI6D, 
               other = H1GI6E, hispanic = H1GI4)

```

### SES variables

We use family income and adolescent reported parental education. Income will be further log-transformed in downstream analysis.

```{r}
ses <- Wave1 |> select(AID, PA55, H1RM1, H1RF1) |>
       mutate(PA55 = na_if(PA55, 9996),
              H1RM1 = if_else(H1RM1 %in% c(96:99, 11, 12), NA, H1RM1),
              H1RF1 = if_else(H1RF1 %in% c(96:99, 11, 12), NA, H1RF1)) |> 
       rename(income = PA55, edu_mum = H1RM1, edu_dad = H1RF1) |> 
       mutate(edu_mum = dplyr::recode(edu_mum,'3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0),
              edu_dad = dplyr::recode(edu_dad, '3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0),
              #income = replace_na(income, median(income, na.rm = T)),
              income = log(income + 1)) |> 
      rowwise() |> 
      mutate(par_ed = mean(c(edu_mum, edu_dad), na.rm = T)) |> 
      ungroup()

ses <- ses |> select(AID, income, par_ed)

```

### Supportive Parenting

```{r}
parent <- Wave1 |> select(AID, H1WP9, H1WP10, H1PF1, H1PF2, 
                            H1PF3, H1PF4, H1PF5, H1WP13, 
                            H1WP14, H1PF23, H1PF24, H1PF25) |> 
            mutate(across(H1WP9:H1PF25, ~ replace(.x, .x > 5, NA)),
                   across(starts_with("H1PF"), ~ 6 - .x),
                   across(H1WP9:H1PF25, ~ as.vector(scale(.x)))) |> 
            rowwise() |> 
            mutate(parent_mum = if_else(sum(is.na(c(H1WP9, H1WP10, H1PF1, 
                                        H1PF2, H1PF3, H1PF4, H1PF5))) <= 3,
                                        mean(c(H1WP9, H1WP10, H1PF1, 
                                        H1PF2, H1PF3, H1PF4, H1PF5), na.rm = T), NA),
                   parent_dad = if_else(sum(is.na(c(H1WP13, H1WP14, H1PF23,
                                        H1PF24, H1PF25))) <= 2,
                                        mean(c(H1WP13, H1WP14, H1PF23,
                                        H1PF24, H1PF25), na.rm = T), NA),
                   parent = mean(c(parent_mum, parent_dad), na.rm = T),
                   highest_parent = max(c(parent_mum, parent_dad), na.rm = T),
                   highest_parent = replace(highest_parent, highest_parent == -Inf, NA)) |> 
            ungroup() |> 
            mutate(single_par = as.numeric(
                                           (is.na(parent_mum)&!is.na(parent_dad))|
                                           (!is.na(parent_mum)&is.na(parent_dad))
                                           ) - .5)
cor(parent$highest_parent, parent$parent, use = "pairwise")
parent |> select(parent, highest_parent) |> skimr::skim()
parent <- parent |> select(AID, parent, single_par)
```

### Delinquent peer affiliation

```{r}
del_peer <- Wave1 |> select(AID, H1TO29, H1TO33, H1TO9, H1DS14) |> 
            mutate(across(H1TO29:H1DS14, ~ replace(.x, .x > 5, NA)),
                   across(H1TO29:H1DS14, ~ as.vector(scale(.x)))) |> 
            rowwise() |> 
            mutate(peer = if_else(sum(is.na(c(H1TO29, H1TO33, H1TO9, H1DS14))) <= 2,
                                  mean(c(H1TO29, H1TO33, H1TO9, H1DS14), na.rm = T),NA)) |> 
            ungroup() 

del_peer <- del_peer |> select(AID, peer)
```

### School connectedness

```{r}
school_con <- Wave1 |> select(AID, H1ED19, H1ED20, H1ED22, H1ED23,
                              H1ED24, H1PR2) |> 
              mutate(across(H1ED19:H1PR2, ~ replace(.x, .x > 5, NA)),
                     across(starts_with("H1ED"), ~ 6 - .x),
                     across(H1ED19:H1PR2, ~ as.vector(scale(.x)))) |> 
              rowwise() |> 
              mutate(school = if_else(sum(is.na(c(H1ED19, H1ED20, H1ED22, H1ED23,
                                      H1ED24, H1PR2))) <= 3,
                                      mean(c(H1ED19, H1ED20, H1ED22, H1ED23,
                                      H1ED24, H1PR2), na.rm = T), NA)) |> 
              ungroup() 

school_con <- school_con |> select(AID, school)
```


### Neighborhood connectedness

```{r}
neigh_con <- Wave1 |> select(AID, H1NB1, H1NB2, H1NB3, H1NB5,
                              H1NB6, H1NB7) |> 
              mutate(across(H1NB1:H1NB7, ~ replace(.x, .x > 5, NA)),
                     across(H1NB1:H1NB3, ~ 3 - .x),
                     H1NB7 = 6 - H1NB7,
                     across(H1NB1:H1NB7, ~ as.vector(scale(.x)))) |> 
              rowwise() |> 
              mutate(neigh = if_else(sum(is.na(c(H1NB1, H1NB2, H1NB3, H1NB5,
                                     H1NB6, H1NB7))) <= 3,
                                     mean(c(H1NB1, H1NB2, H1NB3, H1NB5,
                                     H1NB6, H1NB7), na.rm = T), NA)) |> 
              ungroup()  

neigh_con <- neigh_con |> select(AID, neigh)

```


### EXT PGS (residualized on PCs)

```{r}
pgs_ext <- pgs_pc |> mutate(
  EXT_PGS = lm(EXT_PGS ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + 
                         PC8 + PC9 + PC10 + PC11 + PC12 + PC13 + PC14 +
                         PC15 + PC16 + PC17 + PC18 + PC19 + PC20,
               data = pgs_pc, na.action = "na.exclude") |>  residuals(),
  EXT_PGS = scale(EXT_PGS)|> as.vector()) |> 
  select(AID, FAMID, EXT_PGS, ANCESTRY)

```

### Trajectory

```{r}
traj <- read_csv(here::here(path_data,"Sasia_GMM_results.csv")) |> 
        rename(asb = asb_class, sub = sub_class3) |> 
        mutate(asb = case_match(asb,  1 ~ 2, 2 ~ 4, 3 ~ 1, 4 ~ 3),
               sub = case_match(sub,  1 ~ 1, 2 ~ 3, 3 ~ 2),
               AID = as.character(AID))

```


## Join Data

```{r}
data <- data |> 
          left_join(race) |> 
          left_join(pgs_ext) |> 
          left_join(parent) |> 
          left_join(del_peer) |> 
          left_join(school_con) |> 
          left_join(neigh_con) |> 
          left_join(age) |> 
          left_join(sex) |> 
          left_join(traj) |> 
          left_join(ses) 
```


## Select one from each family per ancestry

```{r}
set.seed(12345)

data <- data |> filter(!is.na(ANCESTRY)) |> group_by(FAMID, ANCESTRY) |> 
        slice_sample(n = 1) |> ungroup()

data |> skimr::skim()
data$ANCESTRY |> table()

```

## Preprocess

### Standardize

```{r}
data |> names()
data <- data |> group_by(ANCESTRY) |> 
  mutate(
         EXT_PGS = scale(EXT_PGS) |> as.vector(),
         parent = scale(parent) |> as.vector(),
         peer = scale(peer) |> as.vector(),
         school = scale(school) |> as.vector(),
         neigh = scale(neigh) |> as.vector(),
         age = scale(age, scale = F) |> as.vector(),
         income = scale(income, scale = F) |> as.vector(),
         par_ed = scale(par_ed, scale = F) |> as.vector(),
        ) |> 
  ungroup()

data |> skimr::skim()

```


### Dummy coded trajectories

```{r}
data <- data |> mutate(
                          sub_hi_dum = if_else(sub == 1, 1, 0),
                          sub_lo_dum = if_else(sub == 2, 1, 0), 
                          asb_hi_dum = if_else(asb == 1, 1, 0),
                          asb_mod_dum = if_else(asb == 2, 1, 0),
                          asb_ado_dum = if_else(asb == 3, 1, 0) 
                          )
```


### Interaction terms

```{r}
data <- data |> mutate(
                       pgsparent = EXT_PGS*parent,
                       pgspeer = EXT_PGS*peer,
                       pgsschool = EXT_PGS*school,
                       pgsneighbor = EXT_PGS*neigh,
                       pgsage = EXT_PGS*age,
                       pgssex = EXT_PGS*sex,
                       pgsinc = EXT_PGS*income,
                       pgspared = EXT_PGS*par_ed,
                       pgssingpar = EXT_PGS*single_par,
                       pgsafrican = EXT_PGS*white,
                       pgsnative = EXT_PGS*native,
                       pgsasian = EXT_PGS*asian,
                       pgsother = EXT_PGS*other,
                       pgshispanic = EXT_PGS*hispanic
                       )

```


## Write data

```{r}
write.table(data,file = here::here(path_data, "gxe_ext_trajectory.csv"),sep = ",", na = ".",row.names = F,col.names  = F)

names(data) |> cat()
```

## Sensitivity analyses: sum scores by waves

```{r}
sum_scores <- read_csv(here::here(path_data, "ASB_SUB_vars_wave_AllParticipants.csv")) |> 
              mutate(AID = as.character(AID))
```


```{r}
write.table(left_join(data, sum_scores),file = here::here(path_data, "gxe_ext_trajectory_wave.csv"),sep = ",", na = ".",row.names = F,col.names  = F)

names(left_join(data, sum_scores)) |> cat()
```

## Descriptives


### Create data for descriptives

```{r}
descrip <- Wave1 |> left_join(Wave2) |> left_join(Wave3) |> 
  left_join(Wave4) |> left_join(Wave5)|> 
  mutate(H1GI1Y = na_if(H1GI1Y, 96),
  H1GI1M = na_if(H1GI1M, 96),
  H2GI1Y = na_if(H1GI1Y, 98),
  H2GI1M = na_if(H1GI1M, 98),
  age1 = (IYEAR - H1GI1Y) + (IMONTH - H1GI1M)/12,
  age2 = (IYEAR2 - H2GI1Y) + (IMONTH2 - H2GI1M)/12,
  age3 = (IYEAR3 - H3OD1Y) + (IMONTH3 - H3OD1M)/12,
  age4 = (IYEAR4 - H4OD1Y) + (IMONTH4 - H4OD1M)/12,
  age5 = (IYEAR5 - H5OD1Y) + (IMONTH5 - H5OD1M)/12,
  income = na_if(PA55, 9996),
  H1RM1 = if_else(H1RM1 %in% c(96:99, 11, 12), NA, H1RM1),
  H1RF1 = if_else(H1RF1 %in% c(96:99, 11, 12), NA, H1RF1),
  edu_mum = dplyr::recode(H1RM1,'3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0),
              edu_dad = dplyr::recode(H1RF1, '3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0)
  ) |> rowwise() |> mutate(par_ed = mean(c(edu_mum, edu_dad), na.rm = T)) |> ungroup() |> left_join(sex) |> left_join(race) |> left_join(parent) |> left_join(sum_scores) |> 
    left_join(pgs_ext) |> left_join(del_peer) |> 
    left_join(school_con) |> left_join(neigh_con) |> 
  select(AID, ANCESTRY, sex, age1:age5, white:hispanic, income, edu_mum, edu_dad, par_ed, single_par, w1asb_s:w5sub_s, EXT_PGS, parent, peer, school, neigh) |> 
  mutate(ANCESTRY = factor(ANCESTRY),
         sex = factor(sex, c(-.5, .5), c("male","female")),
         across(white:hispanic, ~factor(.x)),
         across(edu_mum:single_par, ~factor(.x)),
)

eur <- data |> filter(ANCESTRY == 0 & white == 1 & hispanic == 0) |> 
       select(AID) |> left_join(descrip)

afr <- data |> filter(ANCESTRY == 1 & african == 1 & hispanic == 0) |> 
       select(AID) |> left_join(descrip)

asi <- data |> filter(ANCESTRY == 2 & asian == 1 & hispanic == 0) |> 
       select(AID) |> left_join(descrip)

his <- data |> filter(ANCESTRY == 3 & hispanic == 1) |> 
       select(AID) |> left_join(descrip)

full <- eur |> bind_rows(afr) |> 
  bind_rows(asi) |> bind_rows(his)
```


### Skim

```{r}
descrip |> skimr::skim()
eur |> skimr::skim()
eur$edu_mum |> table()
eur$edu_dad |> table()

afr |> skimr::skim()
afr$edu_mum |> table()
afr$edu_dad |> table()

asi |> skimr::skim()
asi$edu_mum |> table()
asi$edu_dad |> table()

his |> skimr::skim()
his$edu_mum |> table()
his$edu_dad |> table()

```

### Diff test

```{r}
chisq.test(full$sex, full$ANCESTRY)
chisq.test(full$edu_mum, full$ANCESTRY)
chisq.test(full$edu_dad, full$ANCESTRY)

lm(age1~ANCESTRY, full) |> car::Anova(type = 3)
lm(age2~ANCESTRY, full) |> car::Anova(type = 3)
lm(age3~ANCESTRY, full) |> car::Anova(type = 3)
lm(age4~ANCESTRY, full) |> car::Anova(type = 3)
lm(age5~ANCESTRY, full) |> car::Anova(type = 3)

lm(income~ANCESTRY, full) |> car::Anova(type = 3)

lm(w1sub_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w2sub_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w3sub_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w4sub_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w5sub_s~ANCESTRY, full) |> car::Anova(type = 3)

lm(w1asb_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w2asb_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w3asb_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w4asb_s~ANCESTRY, full) |> car::Anova(type = 3)
lm(w5asb_s~ANCESTRY, full) |> car::Anova(type = 3)

lm(parent~ANCESTRY, full) |> car::Anova(type = 3)
lm(peer~ANCESTRY, full) |> car::Anova(type = 3)
lm(school~ANCESTRY, full) |> car::Anova(type = 3)
lm(neigh~ANCESTRY, full) |> car::Anova(type = 3)

```

### Correlations

```{r}
apaTables::apa.cor.table(eur |> select(w1sub_s:w5sub_s, w1asb_s:w5asb_s, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par) |>
                                mutate(income = log(income + 1), 
                                       across(everything(), ~as.numeric(.x))) |> 
                                rowwise() |> mutate(sub = mean(c(w1sub_s, w2sub_s, w3sub_s, 
                                                                 w4sub_s, w5sub_s), na.rm = T),
                                                    asb = mean(c(w1asb_s, w2asb_s, w3asb_s, 
                                                                 w4asb_s, w5asb_s), na.rm = T)) |>
                                ungroup() |> select(sub,asb, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par), 
                         filename = "cor_eur.doc")

apaTables::apa.cor.table(afr |> select(w1sub_s:w5sub_s, w1asb_s:w5asb_s, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par) |>
                                mutate(income = log(income + 1), 
                                       across(everything(), ~as.numeric(.x))) |> 
                                rowwise() |> mutate(sub = mean(c(w1sub_s, w2sub_s, w3sub_s, 
                                                                 w4sub_s, w5sub_s), na.rm = T),
                                                    asb = mean(c(w1asb_s, w2asb_s, w3asb_s, 
                                                                 w4asb_s, w5asb_s), na.rm = T)) |>
                                ungroup() |> select(sub,asb, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par), 
                         filename = "cor_afr.doc")

apaTables::apa.cor.table(asi |> select(w1sub_s:w5sub_s, w1asb_s:w5asb_s, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par) |>
                                mutate(income = log(income + 1), 
                                       across(everything(), ~as.numeric(.x))) |> 
                                rowwise() |> mutate(sub = mean(c(w1sub_s, w2sub_s, w3sub_s, 
                                                                 w4sub_s, w5sub_s), na.rm = T),
                                                    asb = mean(c(w1asb_s, w2asb_s, w3asb_s, 
                                                                 w4asb_s, w5asb_s), na.rm = T)) |>
                                ungroup() |> select(sub,asb, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par), 
                         filename = "cor_asi.doc")

apaTables::apa.cor.table(his |> select(w1sub_s:w5sub_s, w1asb_s:w5asb_s, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par) |>
                                mutate(income = log(income + 1), 
                                       across(everything(), ~as.numeric(.x))) |> 
                                rowwise() |> mutate(sub = mean(c(w1sub_s, w2sub_s, w3sub_s, 
                                                                 w4sub_s, w5sub_s), na.rm = T),
                                                    asb = mean(c(w1asb_s, w2asb_s, w3asb_s, 
                                                                 w4asb_s, w5asb_s), na.rm = T)) |>
                                ungroup() |> select(sub,asb, EXT_PGS, parent, peer, 
                                       school, neigh, sex, age1, income, par_ed, single_par), 
                         filename = "cor_his.doc")

```


## Missing Data

### GMM

```{r}
mda_gmm <- Wave1 |> select(AID) |> left_join(sum_scores) |> 
                    rowwise() |> mutate(sub = mean(c(w1sub_s, w2sub_s, w3sub_s, 
                                                     w4sub_s, w5sub_s), na.rm = T),
                                        asb = mean(c(w1asb_s, w2asb_s, w3asb_s, 
                                                     w4asb_s, w5asb_s), na.rm = T)) |>
                                ungroup()

mda_gmm_plot <- mda_gmm |>  rename("Wave 1 SUB" = w1sub_s, 
                           "Wave 2 SUB" = w2sub_s, 
                           "Wave 3 SUB" = w3sub_s, 
                           "Wave 4 SUB" = w4sub_s, 
                           "Wave 5 SUB" = w5sub_s,
                           "Wave 1 ASB" = w1asb_s, 
                           "Wave 2 ASB" = w2asb_s, 
                           "Wave 3 ASB" = w3asb_s, 
                           "Wave 4 ASB" = w4asb_s, 
                           "Wave 5 ASB" = w5asb_s,
                           "Average SUB" = sub, 
                           "Average ASB" = asb)
mda_gmm_plot |>  select("Wave 1 SUB":"Wave 5 SUB") |> mice::md.pattern(rotate.names = T)
mda_gmm_plot |>  select("Wave 1 ASB":"Wave 5 ASB") |> mice::md.pattern(rotate.names = T)

mda_gmm |> select(w1sub_s:w5sub_s) |> naniar::miss_var_summary()
mda_gmm_sub <-mda_gmm |> select(w1sub_s:w5sub_s) |> naniar::bind_shadow()

lm(w1sub_s ~ w2sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w1sub_s ~ w3sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w1sub_s ~ w4sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w1sub_s ~ w5sub_s_NA, data = mda_gmm_sub) |> summary()

lm(w2sub_s ~ w1sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w2sub_s ~ w3sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w2sub_s ~ w4sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w2sub_s ~ w5sub_s_NA, data = mda_gmm_sub) |> summary()

lm(w3sub_s ~ w1sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w3sub_s ~ w2sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w3sub_s ~ w4sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w3sub_s ~ w5sub_s_NA, data = mda_gmm_sub) |> summary()

lm(w4sub_s ~ w1sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w4sub_s ~ w2sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w4sub_s ~ w3sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w4sub_s ~ w5sub_s_NA, data = mda_gmm_sub) |> summary()

lm(w5sub_s ~ w1sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w5sub_s ~ w2sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w5sub_s ~ w3sub_s_NA, data = mda_gmm_sub) |> summary()
lm(w5sub_s ~ w4sub_s_NA, data = mda_gmm_sub) |> summary()


mda_gmm |> select(w1asb_s:w5asb_s) |> naniar::miss_var_summary()
mda_gmm_asb <-mda_gmm |> select(w1asb_s:w5asb_s) |> naniar::bind_shadow()

lm(w1asb_s ~ w2asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w1asb_s ~ w3asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w1asb_s ~ w4asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w1asb_s ~ w5asb_s_NA, data = mda_gmm_asb) |> summary()

lm(w2asb_s ~ w1asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w2asb_s ~ w3asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w2asb_s ~ w4asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w2asb_s ~ w5asb_s_NA, data = mda_gmm_asb) |> summary()

lm(w3asb_s ~ w1asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w3asb_s ~ w2asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w3asb_s ~ w4asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w3asb_s ~ w5asb_s_NA, data = mda_gmm_asb) |> summary()

lm(w4asb_s ~ w1asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w4asb_s ~ w2asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w4asb_s ~ w3asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w4asb_s ~ w5asb_s_NA, data = mda_gmm_asb) |> summary()

lm(w5asb_s ~ w1asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w5asb_s ~ w2asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w5asb_s ~ w3asb_s_NA, data = mda_gmm_asb) |> summary()
lm(w5asb_s ~ w4asb_s_NA, data = mda_gmm_asb) |> summary()

```


### G-E
```{r}
mda <- data |> select(white:hispanic, EXT_PGS:par_ed)

mda_eur <- mda |> filter(ANCESTRY == 0 & white == 1 & hispanic == 0) 
mda_afr <- mda |> filter(ANCESTRY == 1 & african == 1 & hispanic == 0) 
mda_asi <- mda |> filter(ANCESTRY == 2 & asian == 1 & hispanic == 0) 
mda_his <- mda |> filter(ANCESTRY == 3 & hispanic == 1) 

mda_full <- mda_eur |> bind_rows(mda_afr) |> bind_rows(mda_asi) |> bind_rows(mda_his)

mda_full |> mutate(
                   ANCESTRY = factor(ANCESTRY, 0:3, c("EUR", "AFR", "ASI", "HIS"))
                   ) |> 
            select(-c(white:hispanic)) |> 
  naniar::gg_miss_fct(fct = ANCESTRY)

mda_full |> mutate(
                   ANCESTRY = factor(ANCESTRY, 0:3, c("EUR", "AFR", "ASI", "HIS"))
                   ) |> 
            select(-c(white:hispanic)) |> 
            group_by(ANCESTRY) |> 
            naniar::miss_var_summary() |> print(n = 100)

mda_full |> mice::md.pattern(rotate.names = T)

mda_eur <- mda_eur |> naniar::bind_shadow()
mda_afr <- mda_afr |> naniar::bind_shadow()
mda_asi <- mda_asi |> naniar::bind_shadow()
mda_his <- mda_his |> naniar::bind_shadow()

lm(EXT_PGS ~ par_ed_NA, mda_eur) |> summary()
lm(parent ~ par_ed_NA, mda_eur) |> summary()
lm(peer ~ par_ed_NA, mda_eur) |> summary()
lm(school ~ par_ed_NA, mda_eur) |> summary()
lm(neigh ~ par_ed_NA, mda_eur) |> summary()
lm(age ~ par_ed_NA, mda_eur) |> summary()
glm(sex ~ par_ed_NA, family = binomial("logit"), mda_eur |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(income ~ par_ed_NA, mda_eur) |> summary()
glm(single_par ~ par_ed_NA,  family = binomial("logit"), mda_eur |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ income_NA, mda_eur) |> summary()
lm(parent ~ income_NA, mda_eur) |> summary()
lm(peer ~ income_NA, mda_eur) |> summary()
lm(school ~ income_NA, mda_eur) |> summary()
lm(neigh ~ income_NA, mda_eur) |> summary()
lm(age ~ income_NA, mda_eur) |> summary()
glm(sex ~ income_NA, family = binomial("logit"), mda_eur |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(par_ed ~ income_NA, mda_eur) |> summary()
glm(single_par ~ income_NA,  family = binomial("logit"), mda_eur |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ par_ed_NA, mda_afr) |> summary()
lm(parent ~ par_ed_NA, mda_afr) |> summary()
lm(peer ~ par_ed_NA, mda_afr) |> summary()
lm(school ~ par_ed_NA, mda_afr) |> summary()
lm(neigh ~ par_ed_NA, mda_afr) |> summary()
lm(age ~ par_ed_NA, mda_afr) |> summary()
glm(sex ~ par_ed_NA, family = binomial("logit"), mda_afr |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(income ~ par_ed_NA, mda_afr) |> summary()
glm(single_par ~ par_ed_NA,  family = binomial("logit"), mda_afr |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ income_NA, mda_afr) |> summary()
lm(parent ~ income_NA, mda_afr) |> summary()
lm(peer ~ income_NA, mda_afr) |> summary()
lm(school ~ income_NA, mda_afr) |> summary()
lm(neigh ~ income_NA, mda_afr) |> summary()
lm(age ~ income_NA, mda_afr) |> summary()
glm(sex ~ income_NA, family = binomial("logit"), mda_afr |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(par_ed ~ income_NA, mda_afr) |> summary()
glm(single_par ~ income_NA,  family = binomial("logit"), mda_afr |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ par_ed_NA, mda_asi) |> summary()
lm(parent ~ par_ed_NA, mda_asi) |> summary()
lm(peer ~ par_ed_NA, mda_asi) |> summary()
lm(school ~ par_ed_NA, mda_asi) |> summary()
lm(neigh ~ par_ed_NA, mda_asi) |> summary()
lm(age ~ par_ed_NA, mda_asi) |> summary()
glm(sex ~ par_ed_NA, family = binomial("logit"), mda_asi |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(income ~ par_ed_NA, mda_asi) |> summary()
glm(single_par ~ par_ed_NA,  family = binomial("logit"), mda_asi |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ income_NA, mda_asi) |> summary()
lm(parent ~ income_NA, mda_asi) |> summary()
lm(peer ~ income_NA, mda_asi) |> summary()
lm(school ~ income_NA, mda_asi) |> summary()
lm(neigh ~ income_NA, mda_asi) |> summary()
lm(age ~ income_NA, mda_asi) |> summary()
glm(sex ~ income_NA, family = binomial("logit"), mda_asi |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(par_ed ~ income_NA, mda_asi) |> summary()
glm(single_par ~ income_NA,  family = binomial("logit"), mda_asi |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ par_ed_NA, mda_his) |> summary()
lm(parent ~ par_ed_NA, mda_his) |> summary()
lm(peer ~ par_ed_NA, mda_his) |> summary()
lm(school ~ par_ed_NA, mda_his) |> summary()
lm(neigh ~ par_ed_NA, mda_his) |> summary()
lm(age ~ par_ed_NA, mda_his) |> summary()
glm(sex ~ par_ed_NA, family = binomial("logit"), mda_his |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(income ~ par_ed_NA, mda_his) |> summary()
glm(single_par ~ par_ed_NA,  family = binomial("logit"), mda_his |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()

lm(EXT_PGS ~ income_NA, mda_his) |> summary()
lm(parent ~ income_NA, mda_his) |> summary()
lm(peer ~ income_NA, mda_his) |> summary()
lm(school ~ income_NA, mda_his) |> summary()
lm(neigh ~ income_NA, mda_his) |> summary()
lm(age ~ income_NA, mda_his) |> summary()
glm(sex ~ income_NA, family = binomial("logit"), mda_his |> mutate(sex = case_match(sex, -.5 ~ 0, .5 ~ 1))) |> summary()
lm(par_ed ~ income_NA, mda_his) |> summary()
glm(single_par ~ income_NA,  family = binomial("logit"), mda_his |> mutate(single_par = case_match(single_par, -.5 ~ 0, .5 ~ 1))) |> summary()
```

### GMM in G-E sample

```{r}

eur |> bind_rows(afr) |> bind_rows(asi) |> bind_rows(his) |> select(AID) |> left_join(mda_gmm) |>  select(w1sub_s:w5sub_s) |> mice::md.pattern(rotate.names = T)

eur |> bind_rows(afr) |> bind_rows(asi) |> bind_rows(his)  |> select(AID) |> left_join(mda_gmm) |>  select(w1sub_s:w5sub_s) |> naniar::miss_case_table()

eur |> bind_rows(afr) |> bind_rows(asi) |> bind_rows(his)  |> select(AID) |> left_join(mda_gmm) |>  select(w1asb_s:w5asb_s) |> mice::md.pattern(rotate.names = T)

eur |> bind_rows(afr) |> bind_rows(asi) |> bind_rows(his)  |> select(AID) |> left_join(mda_gmm) |>  select(w1asb_s:w5asb_s) |> naniar::miss_case_table()

```


