---
title: "Add Health Data Preparation for G-E Interplay on EXT Trajectories"
author: "LiChen Dong"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---


## Setup

```{r}
#| output: false

options(conflicts.policy = "depends.ok")
library(tidyverse)
library(lavaan)

path_data <- "../data_local"
```


## Read in data 

### Add health in-home interview data

```{r}
load(here::here(path_data, "add_health_core.RData"))
```

### Genetic data ID - Phenotypic data ID linking file

```{r}
gid <- data.table::fread(here::here(path_data, "AddHealth dbGap ID Link File.dat")) |> as_tibble() |> mutate(AID = as.character(AID))
```

### Genetic ancestry

```{r}
ancestry <- foreign::read.xport(here::here(path_data, "PGS_AH1.xpt")) |> select(AID, ANCESTRY) |>  mutate(AID = as.character(AID))

```

### Polygenic score and genetic PCs

```{r}
ext_pgs <- read_csv(here::here(path_data, "8.14.24_EXT_PGS_PRSCS.csv")) |>
          mutate(EXT_PGS = scale(EXT_PGS_PRSCS) |> as.vector(),
                 AID = as.character(AID)) |> 
          select(-EXT_PGS_PRSCS)

genetic_pc <- data.table::fread(here::here(path_data, "ADDHPCA"))
names(genetic_pc) <- c("GFAM","GID",paste0("PC",1:20))

pgs_pc <- left_join(gid, ext_pgs) |>
           left_join(ancestry) |> 
           left_join(genetic_pc) |> 
           mutate(AID = as.character(AID))
```

## Creating variables

### AID

```{r}
data <- Wave1 |> select(AID)
```

### Participant Biological sex

```{r}
sex <- Wave1 |> select(AID, BIO_SEX) |> 
       rename(sex = BIO_SEX) |> 
       mutate(sex = replace(sex, sex > 5, NA),
              sex = case_match(sex, 1 ~ -.5, 2 ~ .5))
```

### Age

```{r}
age <- Wave1 |> select(AID, H1GI1M, H1GI1Y, IMONTH, IYEAR) |> 
                mutate(H1GI1M = replace(H1GI1M, H1GI1M == 96, NA), 
                       H1GI1Y = replace(H1GI1Y, H1GI1Y == 96, NA)) |> 
                mutate(age = IYEAR - H1GI1Y + (IMONTH - H1GI1M)/12) |> 
                select(-H1GI1M, -H1GI1Y, -IMONTH, -IYEAR) 

```

### Race/ethnicity

```{r}
race <- Wave1 |> select(AID, H1GI6A, H1GI6B, H1GI6C, H1GI6D, H1GI6E, H1GI4) |> 
        mutate(across(H1GI6A:H1GI4, ~ replace(.x, .x > 5, NA))) |> 
        rename(white = H1GI6A, african = H1GI6B, native = H1GI6C, asian = H1GI6D, 
               other = H1GI6E, hispanic = H1GI4)

```

### SES variables

We use family income and adolescent reported parental education. Income will be further log-transformed in downstream analysis.

```{r}
ses <- Wave1 |> select(AID, PA55, H1RM1, H1RF1) |>
       mutate(PA55 = na_if(PA55, 9996),
              H1RM1 = if_else(H1RM1 %in% c(96:99, 11, 12), NA, H1RM1),
              H1RF1 = if_else(H1RF1 %in% c(96:99, 11, 12), NA, H1RF1)) |> 
       rename(income = PA55, edu_mum = H1RM1, edu_dad = H1RF1) |> 
       mutate(edu_mum = dplyr::recode(edu_mum,'3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0),
              edu_dad = dplyr::recode(edu_dad, '3' = 3,
                        '4' = 3,
                        '5' = 3,
                        '6' = 4,
                        '7' = 4,
                        '8' = 5,
                        '9' = 6,
                        '10' = 0),
              #income = replace_na(income, median(income, na.rm = T)),
              income = log(income + 1)) |> 
      rowwise() |> 
      mutate(par_ed = mean(c(edu_mum, edu_dad), na.rm = T)) |> 
      ungroup()

ses <- ses |> select(AID, income, par_ed)

```

### Supportive Parenting

```{r}
parent <- Wave1 |> select(AID, H1WP9, H1WP10, H1PF1, H1PF2, 
                            H1PF3, H1PF4, H1PF5, H1WP13, 
                            H1WP14, H1PF23, H1PF24, H1PF25) |> 
            mutate(across(H1WP9:H1PF25, ~ replace(.x, .x > 5, NA)),
                   across(starts_with("H1PF"), ~ 6 - .x),
                   across(H1WP9:H1PF25, ~ as.vector(scale(.x)))) |> 
            rowwise() |> 
            mutate(parent_mum = if_else(sum(is.na(c(H1WP9, H1WP10, H1PF1, 
                                        H1PF2, H1PF3, H1PF4, H1PF5))) <= 3,
                                        mean(c(H1WP9, H1WP10, H1PF1, 
                                        H1PF2, H1PF3, H1PF4, H1PF5), na.rm = T), NA),
                   parent_dad = if_else(sum(is.na(c(H1WP13, H1WP14, H1PF23,
                                        H1PF24, H1PF25))) <= 2,
                                        mean(c(H1WP13, H1WP14, H1PF23,
                                        H1PF24, H1PF25), na.rm = T), NA),
                   parent = mean(c(parent_mum, parent_dad), na.rm = T)) |> 
            ungroup() |> 
            mutate(single_par = as.numeric(
                                           (is.na(parent_mum)&!is.na(parent_dad))|
                                           (!is.na(parent_mum)&is.na(parent_dad))
                                           ) - .5)

parent <- parent |> select(AID, parent, single_par)
```

### Delinquent peer affiliation

```{r}
del_peer <- Wave1 |> select(AID, H1TO29, H1TO33, H1TO9, H1DS14) |> 
            mutate(across(H1TO29:H1DS14, ~ replace(.x, .x > 5, NA)),
                   across(H1TO29:H1DS14, ~ as.vector(scale(.x)))) |> 
            rowwise() |> 
            mutate(peer = if_else(sum(is.na(c(H1TO29, H1TO33, H1TO9, H1DS14))) <= 2,
                                  mean(c(H1TO29, H1TO33, H1TO9, H1DS14), na.rm = T),NA)) |> 
            ungroup() 

del_peer <- del_peer |> select(AID, peer)
```

### School connectedness

```{r}
school_con <- Wave1 |> select(AID, H1ED19, H1ED20, H1ED22, H1ED23,
                              H1ED24, H1PR2) |> 
              mutate(across(H1ED19:H1PR2, ~ replace(.x, .x > 5, NA)),
                     across(starts_with("H1ED"), ~ 6 - .x),
                     across(H1ED19:H1PR2, ~ as.vector(scale(.x)))) |> 
              rowwise() |> 
              mutate(school = if_else(sum(is.na(c(H1ED19, H1ED20, H1ED22, H1ED23,
                                      H1ED24, H1PR2))) <= 3,
                                      mean(c(H1ED19, H1ED20, H1ED22, H1ED23,
                                      H1ED24, H1PR2), na.rm = T), NA)) |> 
              ungroup() 

school_con <- school_con |> select(AID, school)
```


### Neighborhood connectedness

```{r}
neigh_con <- Wave1 |> select(AID, H1NB1, H1NB2, H1NB3, H1NB5,
                              H1NB6, H1NB7) |> 
              mutate(across(H1NB1:H1NB7, ~ replace(.x, .x > 5, NA)),
                     across(H1NB1:H1NB3, ~ 3 - .x),
                     H1NB7 = 6 - H1NB7,
                     across(H1NB1:H1NB7, ~ as.vector(scale(.x)))) |> 
              rowwise() |> 
              mutate(neigh = if_else(sum(is.na(c(H1NB1, H1NB2, H1NB3, H1NB5,
                                     H1NB6, H1NB7))) <= 3,
                                     mean(c(H1NB1, H1NB2, H1NB3, H1NB5,
                                     H1NB6, H1NB7), na.rm = T), NA)) |> 
              ungroup()  

neigh_con <- neigh_con |> select(AID, neigh)

```


### EXT PGS (residualized on PCs)

```{r}
pgs_ext <- pgs_pc |> mutate(
  EXT_PGS = lm(EXT_PGS ~ PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + 
                         PC8 + PC9 + PC10 + PC11 + PC12 + PC13 + PC14 +
                         PC15 + PC16 + PC17 + PC18 + PC19 + PC20,
               data = pgs_pc, na.action = "na.exclude") |>  residuals(),
  EXT_PGS = scale(EXT_PGS)|> as.vector()) |> 
  select(AID, FAMID, EXT_PGS, ANCESTRY)

```

### Trajectory

```{r}
traj <- read_csv(here::here(path_data,"Sasia_GMM_results.csv")) |> 
        rename(asb = asb_class, sub = sub_class3) |> 
        mutate(asb = case_match(asb,  1 ~ 2, 2 ~ 4, 3 ~ 1, 4 ~ 3),
               sub = case_match(sub,  1 ~ 1, 2 ~ 3, 3 ~ 2),
               AID = as.character(AID))

```


## Join Data

```{r}
data <- data |> 
          left_join(race) |> 
          left_join(pgs_ext) |> 
          left_join(parent) |> 
          left_join(del_peer) |> 
          left_join(school_con) |> 
          left_join(neigh_con) |> 
          left_join(age) |> 
          left_join(sex) |> 
          left_join(traj) |> 
          left_join(ses) 
```


## Select one from each family per ancestry

```{r}
set.seed(12345)

data <- data |> filter(!is.na(ANCESTRY)) |> group_by(FAMID, ANCESTRY) |> 
        slice_sample(n = 1) |> ungroup()

data |> skimr::skim()
data$ANCESTRY |> table()

```

## Preprocess

### Standardize

```{r}
data |> names()
data <- data |> group_by(ANCESTRY) |> 
  mutate(
         EXT_PGS = scale(EXT_PGS) |> as.vector(),
         parent = scale(parent) |> as.vector(),
         peer = scale(peer) |> as.vector(),
         school = scale(school) |> as.vector(),
         neigh = scale(neigh) |> as.vector(),
         age = scale(age, scale = F) |> as.vector(),
         income = scale(income, scale = F) |> as.vector(),
         par_ed = scale(par_ed, scale = F) |> as.vector(),
        ) |> 
  ungroup()

data |> skimr::skim()

```


### Dummy coded trajectories

```{r}
data <- data |> mutate(
                          sub_hi_dum = if_else(sub == 1, 1, 0),
                          sub_lo_dum = if_else(sub == 2, 1, 0), 
                          asb_hi_dum = if_else(asb == 1, 1, 0),
                          asb_mod_dum = if_else(asb == 2, 1, 0),
                          asb_ado_dum = if_else(asb == 3, 1, 0) 
                          )
```


### Interaction terms

```{r}
data <- data |> mutate(
                       pgsparent = EXT_PGS*parent,
                       pgspeer = EXT_PGS*peer,
                       pgsschool = EXT_PGS*school,
                       pgsneighbor = EXT_PGS*neigh,
                       pgsage = EXT_PGS*age,
                       pgssex = EXT_PGS*sex,
                       pgsinc = EXT_PGS*income,
                       pgspared = EXT_PGS*par_ed,
                       pgssingpar = EXT_PGS*single_par,
                       pgsafrican = EXT_PGS*white,
                       pgsnative = EXT_PGS*native,
                       pgsasian = EXT_PGS*asian,
                       pgsother = EXT_PGS*other,
                       pgshispanic = EXT_PGS*hispanic
                       )

```


## Write data

```{r}
write.table(data,file = here::here(path_data, "gxe_ext_trajectory.csv"),sep = ",", na = ".",row.names = F,col.names  = F)

names(data) |> cat()
```


## Descriptives

### Create data for descriptives

```{r}
descrip <- data |> select(AID, ANCESTRY, sex, white, african, native, 
                          asian, other, hispanic, single_par, EXT_PGS)
```

### Read in sum scores of ASB and SUB

### Skim


